Generalised gravitational burst generation with Generative Adversarial Networks

### 2020-09-09 Paper_draft/main.tex
addressed comments up to section 3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\bibliographystyle{iopart-num_custom}
\usepackage{xcolor}
\usepackage{pifont}
\usepackage{comment}
\usepackage{acronym}
\usepackage{standalone}
\usepackage{academicons}
\usepackage{scalerel}
\usepackage{layouts}
%\usepackage{import}
\usepackage{subfig}
\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing,angles,quotes,shapes.geometric, arrows,calc}
\usetikzlibrary{svg.path}
\expandafter\let\csname equation*\endcsname\relax 
\expandafter\let\csname endequation*\endcsname\relax 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage[capitalize]{cleveref}

@@

% more details on burst searches
%
\ac{GW} searches for modelled signals so far have used a process called
matched-filtering,~\cite{Owen1998}, where a large template bank of possible
\ac{GW} waveforms are compared to the detector outputs. In order to increase
the chances of detection these template banks must span a large
multi-dimensional parameter space and the templates must be accurate, requiring
significant computational cost~\chris{the accuracy doesn't obviously lead to a
cost increase}. \ac{GW} bursts are un-modelled~\chris{except for the ones you
mentione before? SNe, glicthes, post-merger? }; therefore there are no
templates available and so matched-filtering is unsuitable for the detection of
these signals.  Instead, detection involves distinguishing the signal from
detector noise~\chris{but that's also what matched-filtering does}. Burst
searches look for excess power contained in the time-frequency
domain~\chris{all of them? In any case you have to provide references here to
the pipelines (cWB, x-pipeline, etc... Siong can help)} and rely on the
astrophysical burst waveform appearing in multiple detectors at similar times.
This is only possible if the detector noise is well characterised and the
candidate signal can be differentiated from systematic or environmental
glitches.

% Discuss the family of burst waveforms currently used and why - not in detail, just
% an introduction 
%
\ac{GW} burst detection algorithms~\cite{Klimenko_2008, Aso_2008} are tested
and tuned using modelled waveforms that may or may not have astrophysical
significance but have easy to define parameters and share characteristics of
real bursts that are enough to simulate a \ac{GW} passing between
detectors~\chris{long sentence}. Such waveforms include sine-Gaussians: a
Gaussian modulated sine wave that is characterised by its central frequency and
decay parameter. Bandlimited white-noise bursts: white noise that is contained
within a certain frequency range and ring-downs which mimic the damped
oscillations after a \ac{CBC} merger.~\chris{seems strange to list only 3 of
the 5 that we will be using later on.}

% cut this text if not going to be used
%
\begin{comment}Such waveforms may have
long-duration, short bandwidth (ringdowns), long-duration, large bandwidth
(inspirals) and many algorithms make use of sine-Gaussians: a Gaussian
modulated sine wave that is characterised by it's central frequency and narrow
bandwidth.~\chris{not the time to try to describe the types of burst waveforms.
Also, be careful with satying things like ringdowns have narrow bandwidth. I
know we specify a single frequency but becuase of the short duration, the
signal is broad band. Look at the FFT} This makes it a great tool for
diagnosing LIGOs sensitivity to frequency.~\chris{strange unfinished sentence.} 
\end{comment}

% Introduce ML techniques in GWs
%
With the expectation that there will be many more \ac{GW} detections in the
future, there is a growing need for fast and efficient \ac{GW} analysis methods
to match the rising number of detections. While still in its infancy, \ac{ML}
applications to \ac{GW} analyses have already shown great potential for
overcoming data bottlenecks~\chris{what does that mean?}. \ac{GW} analysis in
areas of detection~\cite{Gabbard2017,Gebhard_2019,Krastev_2020}, glitch
classification~\cite{Bahaadini, George_2018,Razzano_2018} and parameter
estimation~\cite{gabbard2019bayesian, shen2019deterministic,
green2020gravitationalwave}~\chris{unfinished sentence, plus I wouldn't
reference the Shen paper}. Applying these algorithms show similar results to
matched-filtering techniques and have the benefit of being orders of magnitude
faster. \michael{We mention aspects related to detection (pattern matching) but
also metion PE in the same sentence and I think this is a bit
confusing.}~\chris{yes, the matched-filtering comp[arison is only relevent for
searches. The biggest speed gain is for the Bayesian PE. I also think that you
need to reference more of the ML-GW literature here. I've added Mervyn's SNe
paper, Joe's latest paper, our latest Gaussian ?Mixture model paper, 2 CW
papers and the recent ML review paper to the bibligrapohy file but there are
quite a few others. Have a look at the review to steal references. In each case
try to say 1 sentence about what they do if possible.}

% What this paper does on GANs in 1 paragraph - the point of the paper
%
In this work we aim to explore the use of \ac{ML} to generate and interpret
unmodelled \ac{GW} burst waveforms. Using the generative machine learning
model: \acp{GAN}, we train on five classes of known~\chris{are they known? In
what sense?} burst morphologies~\chris{fancy word - what about waveform class?}
in the time domain. Working with the assumption that \acp{GAN} construct smooth
$n$-dimensional vector spaces between their input and output, we can then
explore the space between the five classes to construct new
unmodelled~\chris{hybrid?} waveforms. As all the computationally expensive
processes occur during training, the learned model after training is able to
produce waveforms in fractions of a second and in fact produce waveforms that
are impossible~\chris{never say never (or impossible)} to produce with current
techniques. These new varieties of waveforms can then be used to diagnose
detection algorithms and gain new insight into sources of \ac{GW}
bursts~\chris{yes, they can help diagnose or calibrate or assess the sensitivty
of detection algorithms but also allow us to better train our algorithms to a
broader range of possible signals and therefore enhance our detection ability}.

@@

\begin{figure}
    \begin{minipage}[b]{0.3\linewidth}
        \subfloat[]{\label{fig:perceptron}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure0.pdf}}
        \par
        \vspace{\baselineskip}
        \subfloat[]{\label{fig:tanh_activation}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure1.pdf}}
    \end{minipage}
        \begin{minipage}[b]{0.7\linewidth}
        \subfloat[]{\label{fig:neural_network}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure2.pdf}}
    \end{minipage}
    \caption{Neural Networks (a) A single neuron taking a vector of inputs and
returning a single output based on the weights, bias and activation function
of the network. (b) The hyperbolic tangent used as an activation
function.~\chris{is this the actual activation that you use in the GAN?
otherwise why focus on it alone, you could plot a handful on the same plot} (c)
A fully connected~\chris{you haven't mentioned what a fully connected network
actually is in the text} neural network containing two hidden layers that performs a
mapping of an input vector to a single output. \michael{Equation in the
neuron looks very small}~\chris{I agree, but the equation in (b) is even
smaller. I would also recommed that you make the style of the FC network hidden
and output layer neurons the same as the single perceptron in terms of colours.
You could put $x_i$s in the input layer.}}
\end{figure}

% inroduce the concept of ML
%
Since the birth of programmable computers, people have wondered if machines can
develop true intelligence beyond formal mathematical procedures. Today, \ac{ML}
aims to learn apparent relationships held within the given data or `training
data' in order to make accurate predictions without the need for additional
programming. A common approach in \ac{ML} relies on the computer learning on
past experience to make decisions on future events. \michael{Not sure about the
use of computer here, maybe \textit{relies on a model}?} The hierarchical
nature of this approach means that the machine can build complicated concepts
from simpler ones~\chris{what hierarchical nature? You haven't mentioned
anything about that.}. This branch of \ac{AI} is called deep
learning~\chris{OK, but as long as you explain the hierarchy thing and how it
implies complicated concepts from simpler ones.}.

% Introduce basic neural networks - a perceptron layer
%
Neural networks are the quintessential~\chris{really? quintessential?} \ac{ML}
algorithm that aims to approximate a function. They are built from many single
processing units called neurons. The simpliest Neural Network is the perceptron
layer~\cref{fig:perceptron} which holds a single neuron that takes several real
inputs $x_{1},\ldots, x_{i}$~\chris{and outputs what? also $i$ is the index and not the "final" index}
and maps them to an output according to the linear equation 
%
\begin{align}
\sigma(\sum_i w_i x_i + b),
\label{eqn:neuron}
\end{align}
%
\chris{it's not an equation if it's not equal to something. Please make it
eaqual something.} where $w$ and $b$ are the weights and bias and $\sigma$
denotes the activation function. The weights are numbers which can be thought
as the strength between connections~\chris{what connestions? between what
objects?}. The output of a neuron is called its activation and different
choices of activation function allow the user~\chris{are we users or developers
or what?} to simulate various models. A common example of an activation
function is the hyperbolic tangent~\cref{fig:tanh_activation} that is used when
the outputs of the network must be both positive and negative~\chris{how can an
output be both positive and negative. The language is ambiguous and confusing}
and re-scaled to $[-1,1]$~\chris{at this stage in the explanation the reader
has no idea what an output would look like or why it would be rescaled}. It is
often useful to introduce a bias, $b$, such that the neuron remains inactive
above zero but is active when the sum reaches a defined threshold. This bias is
added before the activation function and it tells us how high the weighted sum
needs to be before the neuron becomes active.~\chris{well kind of, but also not
really. I would advise you to rewrite this (if we keep it) in a minimalist
style just explaining what each item is and not trying to describe what it
does.}

% the cost or loss function
%
The output of a single neuron is defined by~\cref{eqn:neuron} and gives a
prediction $\hat{y}$ that can be compared to the real value $y$~\chris{what the
hell is $y$, it's not been mentioned before} through a loss (also known as a
cost) function. If the loss is non-zero~\chris{no, it doesn't matter whether
the number is above or below or equal to zero, the machine will try to minimise
the number (if negative, it will try to make it more negative)}, the network
must work to minimise this function by updating the weights in the negative
direction of the loss gradient in a process referred to as gradient
descent~\chris{give reference and also make the distinction between this and
the actual method used to do it, back progagation}. This loss function plays a
critical role in how neural networks learn~\chris{OK, but are you going to
expand on that or just leave it there? As far as I can tell from the text it's
the *only* way network's learn so that's pretty critical and already evident}.

% basic network structure
%
A neural network contains many single neurons connected in a layered structure
as shown in~\cref{fig:neural_network}. The activations of the first layer (or
input layer) act as the inputs to the second layer and so on until the output
layer. Multi-layered neural networks have intermediate layers between the input
and output stages dubbed the hidden layers as the computations performed are
not accessible to the user~\chris{well that's not really true. We can access
everything if we want but it depends on what you define as a user. Best to
simply state what they're clalled and no need to give a reason.}. The network
can be thought of as a function F: $\mathbb{R}^N\rightarrow
\mathbb{R}^M$~\chris{does F not have to be in math-mode too? Is is a
mathematical symbol?} reliant on how activations from one layer bring
activations in the next~\chris{yes, kind of, but also reliant on the trained
weights and biases too, so actually reliant on everything}. The system is
analogous to biology, where, some groups of neurons in animal brains cause
certain others to fire~\chris{best to stay away from analogies like these where
the practical specifics can be picked apart}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Convolutional Neural Networks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\acp{CNN} are designed to work with grid-like input structures that exhibit
strong local spatial dependencies. An example would be a two-dimensional image
as each coloured pixel has similar intensities to their near
neighbours~\chris{it's not about that really, it's like you said, it's about
spatial relationships}. The value contained in each pixel can be thought of as
adding an additional dimension or channel to the image~\chris{no, the dimension
and channel are specifically defined. A 2D image has 2 dimensions. An RGB image
has 3 numbers per pixel and these are typcally stored as channels. Channels are
dealt with differently to dimensions}. Although most work with \acp{CNN}
involve image-based data, they can be applied to other spatially adjacent data
types such as time-series and text items. \acp{CNN} are defined by the use of a
convolution operation, a mathematical operation that expresses the amount
overlap between functions. The convolution is applied by shifting one grid-like
structure over the other, drawing out spatially important features between the
two~\chris{yes, but what is being dragged over what? Time to define the
convolutional neuron? Maybe add an equation?}. A strided convolution reduces
the dimensionality of the input while retaining spatial features \michael{How?
This is core part of the architecture of the GAN but is quite vague, I don't
think it's obvious that stride imples moving the filter further. Maybe
something like: \textit{the distance by which the grid is shifted is known as
the stride and increasing it reduces the dimensionality of the
output}}~\chris{why are we talking about strides at this point? Are they
important for the GAN? You haven't mentioned kernel sizes or dropout. are these
equally important?}. This can result in a loss of information along the
borders~\chris{the borders of what?}. To combat this, the input is padded with
zeros at the edges which does not interfere with the convolutional operation.
The output of the convolutional layer is then passed to an activation function.

~\chris{I'm getting the sense that these 2 sections of ANNs and CNNs aren't
really helpful in their current state. They should either be removed OR tidied
up. I don't think they need to be expanded upon but various statements can be
removed and other important things can be added. Imagine reading this with no
knowledge of ML. Do you get informed by this or confused?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GANs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% basic intro to GANs
%
A subset of deep learning that has seen fruitful development in recent years
are \acp{GAN}~\cite{Goodfellow2014}. These unsupervised~\chris{they have
training data don't they? Is that unsupervised?} algorithms learn patterns in a
given training data set using an adversarial process. The generations from
\acp{GAN} are state-of-the-art~\chris{a phrase that will not stand the test of
time} in fields such as high quality image
fidelity~\cite{brock2018large,karras2019analyzing}, text-to-image
translation~\cite{reed2016generative}, and video
prediction~\cite{liang2017dual} as well as time series
generations~\cite{esteban2017realvalued}~\chris{do we still not have any other
GW GAN applications to reference or are we hoping to still be the first?}.

% Basic components of a GAN
%
\acp{GAN} train two competing neural networks, consisting of a discriminator
network that is set up to distinguish between real and fake data and a
generator network that produces synthetic (fake) versions of the real data.
\michael{Is it clear to the reader that the \textit{fake} data is the same as
the synthetic data?} The generator performs a mapping from an input noise
vector $\mathbf{z}$, that is usually sampled from a $n$-dimensional~\chris{be
careful with statments about $n$-dimensions since $n$ is a variable and could
mean the number of latent dimensions or the number of class dimensions or data
dimensions, etc...} Gaussian space known as the latent space, to its
representation of the data, and the discriminator maps its input $\mathbf{x}$ to
a probability that the input came from either the training (real) data or
generator (fake).

% training a GAN
%
During training, the discriminator is given a batch~\chris{you have not
introduced batches yet. Best to do that in the ANN section} of samples that
contains one half real data and one half fake data which it then makes
predictions on.  The loss for the discriminator is calculated by comparing its
predictions to the labelled data through the binary cross-entropy function.
\jordan{note to myself: This is important since eqn 2 is derived from it (thats
where the logs come from)}~\chris{either give a reference to binary cross
entropy or define it here mathematically}. The training process of a \ac{GAN}
alternatively updates the weights of the discriminator and generator based on
information on its competitors loss function~\chris{you haven't fully explained
how the loss is computed in the 2 phases of each training pass. I recommend
that you add that here other wise discussion of generator loss and discrimator
loss makes no sense.}. This loss of the discriminator is used to update the
weights of the generator to produce more realistic samples of the input
distribution, the loss of the generator encourages the discriminator to update
its classification abilities. \acp{GAN} seek to find an equilibrium between the
generator and discriminator in the form of a two-player game that can be
summarised by 
%
\begin{align}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\bibliographystyle{iopart-num_custom}
\usepackage{xcolor}
\usepackage{pifont}
\usepackage{comment}
\usepackage{acronym}
\usepackage{standalone}
\usepackage{academicons}
\usepackage{scalerel}
\usepackage{layouts}
\usepackage{import}
%\usepackage{subfig}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing,angles,quotes,shapes.geometric, arrows,calc}
\usetikzlibrary{svg.path}
\expandafter\let\csname equation*\endcsname\relax 
\expandafter\let\csname endequation*\endcsname\relax 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage[capitalize]{cleveref}

\usepackage{graphicx}
\usetikzlibrary{decorations.pathreplacing,angles,quotes,shapes.geometric, arrows,calc}
\usepackage{wrapfig, blindtext}
\usepackage{pgfplots}
\pgfplotsset{compat=1.16}

@@

% more details on burst searches
%
\ac{GW} searches for modelled signals so far have used a process called
matched-filtering,~\cite{Owen1998}, where a large template bank of possible
\ac{GW} waveforms are compared to the detector outputs. In order to increase
the chances of detection these template banks must span a large
multi-dimensional parameter space requiring
significant computational cost. For \ac{GW} bursts that remain un-modelled; there are no
templates available and so matched-filtering is unsuitable for the detection of
these signals.  Instead, detection algorithms like (cWB, x-pipeline, etc... \jordan{reference}) involve distinguishing the signal from
detector noise by looking for excess power contained in the time-frequency
domain and rely on the
astrophysical burst waveform appearing in multiple detectors at similar times.
This is only possible if the detector noise is well characterised and the
candidate signal can be differentiated from systematic or environmental
glitches.

% Discuss the family of burst waveforms currently used and why - not in detail, just
% an introduction 
%
\ac{GW} burst detection algorithms~\cite{Klimenko_2008, Aso_2008} are tested
and tuned using modelled waveforms that may or may not have astrophysical
significance but have easy to define parameters and share characteristics of
real bursts that are enough to simulate a \ac{GW} passing between
detectors~\chris{long sentence}. Such waveforms include sine-Gaussians: a
Gaussian modulated sine wave that is characterised by its central frequency and
decay parameter. Bandlimited white-noise bursts: white noise that is contained
within a certain frequency range and ring-downs which mimic the damped
oscillations after a \ac{CBC} merger. A Gaussian blip: a short exponential increase the decrease in amplitude and a Binary black hole in spiral. 
% cut this text if not going to be used
%
\begin{comment}Such waveforms may have
long-duration, short bandwidth (ringdowns), long-duration, large bandwidth
(inspirals) and many algorithms make use of sine-Gaussians: a Gaussian
modulated sine wave that is characterised by it's central frequency and narrow
bandwidth.~\chris{not the time to try to describe the types of burst waveforms.
Also, be careful with satying things like ringdowns have narrow bandwidth. I
know we specify a single frequency but becuase of the short duration, the
signal is broad band. Look at the FFT} This makes it a great tool for
diagnosing LIGOs sensitivity to frequency.~\chris{strange unfinished sentence.} 
\end{comment}

% Introduce ML techniques in GWs
%
With the expectation that there will be many more \ac{GW} detections in the
future, there is a growing need for fast and efficient \ac{GW} analysis methods
to match the rising number of detections. While still in its infancy, applying \ac{ML} to \ac{GW} analyses have already shown great potential in
areas of detection~\cite{Gabbard2017,Gebhard_2019,Krastev_2020}, where these techniques have matched the sensitivity of matched filtering for Advanced LIGO and Advanced Virgo gravitational-wave searches. In identifying and classifying detector noise transients or  'glitches"~\cite{Bahaadini, George_2018,Razzano_2018, 2020arXiv200801262G}. In Bayesian parameter
estimation~\cite{gabbard2019bayesian,
green2020gravitationalwave} where \ac{ML} techniques can recover parameters of a \ac{GW} signal significantly faster than standard methods. Long duration signals like continuous \ac{GW} require long observing times and therefore have large amounts of data needing to be processed. Current \ac{ML} approaches~\cite{2020PhRvD.102b2005D, 2019PhRvD.100d4009D, 2020arXiv200708207B} are particularly well suited to dealing with this as once trained the searches can be performed quickly. \jordan{quicker than than what?}

% What this paper does on GANs in 1 paragraph - the point of the paper
%
In this work we aim to explore the use of \ac{ML} to generate and interpret
unmodelled \ac{GW} burst waveforms. Using the generative machine learning
model, \acp{GAN}, we train on five classes of waveforms in the time domain. Working on the assumption that \acp{GAN} construct smooth
 high dimensional vector spaces between their input and output, we can then
explore the space between the five classes to construct new
hybrid waveforms. As all the computationally expensive
processes occur during training, the learned model after training is able to
produce waveforms in fractions of a second and in fact produce waveforms that
are difficult to produce with current
techniques. These new varieties of waveforms can then be used to diagnose
detection algorithms, gain new insight into sources of \ac{GW}
bursts and  allow us to better train our algorithms to a
broader range of possible signals and therefore enhance our detection ability.

@@

% inroduce the concept of ML
%

\begin{figure}[h!]
\begin{subfigure}[b]{0.30\textwidth}
   \begin{subfigure}[b]{1\textwidth}
   	\centering
	\resizebox{\textwidth}{!}{
 	\begin{tikzpicture}		
		    \node[scale=0.3] 		   (x1) at (0,0.8) {$x_1$}; 
		    \node[scale=0.3] 		   (x2) at (0,0.5) {$x_2$};  
		    \node[scale=0.3] 		   (x3) at (0,0.2) {$x_3$};       
    		    \node [draw, circle,scale=0.18,fill=blue!25]          (c3) at (0.7,0.5) {\Large $\sigma(\sum_i w_i x_i + b)$};
    \draw[->] (x1)--(c3);
    \draw[->] (x2)--(c3);
    \draw[->] (x3)--(c3);
     \draw[->] (c3)--(1.3,0.5);
   	\end{tikzpicture}
	}
	\caption{}
	 \label{fig:perceptron}
 \end{subfigure}

   \begin{subfigure}[b]{1\textwidth}
  	\centering
	\resizebox{\textwidth}{!}{
\begin{tikzpicture}
\begin{axis}[
    xmin=-2.5, xmax=2.5,
    ymin=-1.5, ymax=1.5,
    axis lines=center,
    axis on top=true,
    domain=-2.5:2.5,
    ylabel=$\sigma$,
    xlabel=$x$,
    ]

    \addplot [mark=none,draw=green,ultra thick] {tanh(\x)};
    \addplot+[mark=none,draw=blue,domain=-2.5:0,ultra thick] {0};
    \addplot+[mark=none,draw=blue,domain=0:2.5,ultra thick] {x};
    \addplot+[mark=none,red,dotted,domain=-2:0,ultra thick] {0.2*x};
    \addplot+[mark=none,red,dotted,domain=0:2.5,ultra thick] {x};

    %% Add the asymptotes
\end{axis}
\end{tikzpicture}
	}
	\caption{}
	 \label{fig:activations}
\end{subfigure}
\end{subfigure}
	 \begin{subfigure}[b]{0.70\textwidth}
\centering
\resizebox{\textwidth}{!}{
\def\layersep{1.5cm}
\begin{tikzpicture}[shorten >=1pt,draw=black, ->,node distance=\layersep]

    \tikzstyle{every pin edge}=[<-,shorten <=1pt]
    \tikzstyle{neuron}=[circle,draw=black, very thick, fill=blue!25, minimum size=22pt,inner sep=0pt]
    \tikzstyle{input neuron}=[circle,draw=black,very thick,fill=black!25, minimum size=22pt,inner sep=0pt];
    \tikzstyle{output neuron}=[neuron];
    \tikzstyle{hidden neuron}=[circle,draw=black,very thick,fill=blue!25,minimum size=22pt,inner sep=0pt];
    \tikzstyle{annot} = [text width=4em, text centered]

    % Draw the input layer nodes
    \foreach \name / \y in {1,...,5}
    % This is the same as writing \foreach \name / \y in {1/1,2/2,3/3,4/4}
        \node[input neuron] (I-\name) at (0,-\y) {$x_{\name}$};

    % Draw the hidden layer nodes
    \foreach \name / \y in {1,...,3}
        \path[yshift=-1cm]
            node[hidden neuron] (H-\name) at (\layersep,-\y cm) {};

  \foreach \namee / \y in {1,...,3}
        \path[yshift=-1 cm]
            node[hidden neuron] (H1-\namee) at (2*\layersep,-\y cm){};

    % Draw the output layer node
  \node[output neuron,pin={[pin edge={->}]right:}, right of=H1-2] (O) {};
    % Connect every node in the input layer with every node in the
    % hidden layer.
    \foreach \source in {1,...,5}
        \foreach \dest in {1,...,3}
            \path (I-\source) edge (H-\dest);

 \foreach \source in {1,...,3}
        \foreach \dest in {1,...,3}
            \path (H-\source) edge (H1-\dest);

    % Connect every node in the hidden layer with the output layer
    \foreach \source in {1,...,3}
        \path (H1-\source) edge (O);

    % Annotate the layers
    %\node[annot,above of=H-1, node distance=1cm] (hl) {Hidden layers};
   % \node[annot,left of=I-4] at (0) {Input layer};
    %\node[annot,right of=hl] {Output layer};

    \draw[decoration={brace,mirror,raise=15pt},decorate,-]
  (I-1.north) -- node[left=0.8cm,scale=0.7] {Input layer} (I-5.south);

    \draw[decoration={brace,raise=15pt},decorate,-]
  (H-1.west) --  node[above=0.8cm,scale=0.7]{Hidden layers}  (H1-1.east);

    \draw[decoration={brace,raise=15pt},decorate,-]
  (O.west) --  node[above=0.8cm,scale=0.7]{Output layers}  (O.east);

\end{tikzpicture}
}
 \caption{}
 \label{fig:network}
 \end{subfigure}

\caption{Neural Networks (a) A single neuron taking a vector of inputs and
returning a single output based on the weights, bias and activation function
of the network. (b) A selection of activation functions used in this study. The hyperbolic tangent (Green), Rectified linear unit \cite{relu} (Blue) and Leaky rectified linear unit \cite{Maas2013RectifierNI} (Red). (c)
A an example of a neural network containing two hidden layers that performs a
mapping of an input vector to a single output.}
\end{figure}

\ac{ML} aims to learn apparent relationships held within given data or `training
data' in order to make accurate predictions without the need for additional
programming. A common approach in \ac{ML} relies on the model learning on
past experience to make decisions on future events. Deep learning is a branch of \ac{ML} that takes inspiration from biological processes in the brain.

% Introduce basic neural networks - a perceptron layer
%
\begin{comment}
Neural networks are the quintessential~\chris{really? quintessential?} \jordan{It's a perfectly cromulent word https://www.imdb.com/title/tt0701155/} \ac{ML}
algorithm that aims to approximate a function. 
\end{comment}

Artificial neural networks are universal function approximators that are built from many single
processing units called neurons. The simplest neural network is the perceptron
layer~\cref{fig:perceptron} which holds a single neuron that takes several real
inputs $x_{i},\ldots, x_{n}$
and maps them to an \jordan{probability?} output according to the linear function, 
%
\begin{align}
f(x) = \sigma(\sum_i w_i x_i + b),
\label{eqn:neuron}
\end{align}
%
where $w$ and $b$ are the weights and bias and $\sigma$
denotes the activation function. The weights are numbers which can be thought
of as the strength between connected neurons. The output of a neuron is defined by its activation function which controls how the neuron 'fires" depending on its input. Some examples of commonly used activation functions are shown in~\cref{fig:activations}. It is often useful to introduce a bias, $b$, such that the neuron remains inactive
above zero but is active when the sum reaches a defined threshold.

% basic network structure
%
A neural network contains many single neurons connected in a layered structure
as shown in~\cref{fig:network}. The activation's of the first layer (or
input layer) act as the inputs to the second layer and so on until the output
layer. Multi-layered neural networks have intermediate layers between the input
and output stages dubbed the hidden layers.

% the cost or loss function
%
The output of a single neuron is gives a
prediction that can be compared to the real value through a loss (also known as a
cost) function. The network must work to minimise this function by updating the weights in the negative
direction of the loss gradient in a process referred to as gradient
descent \cite{ruder2016overview}. The training process for a single layered network is easy to compute as the weights relate directly to the gradient of the loss function the network is trying to minimise. For deeper architectures, the loss is a complicated function of all the weights in all the layers. The backpropagation \cite{Nielsen1992} algorithm acts over the many paths from node to output. It does so in two phases:

\begin{itemize}
\item Forward phase: For one instance of training, the inputs are fed forward through the network using the current weights and the final output is compared to the training labels. The derivative of the loss function is then computed.
\item Backward phase: This phase learns how the gradient of the loss function changes when the weights are varied. Starting at the output node, the algorithm goes backwards through the network (hence the name). The weights that give the steepest descent to the loss function are saved for the next training instance.  
\end{itemize}
This process of updating the weights is repeated until the loss function reaches convergence or a global minimum. The training data is cycled through in epochs. An epoch is defined when the entire data set is fed through the network. It is impractical to feed the entire data set into the network at once, so the training is split up into smaller more manageable batches. Finally, the number of batches needed to compete one epoch is referred to as the iterations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Convolutional Neural Networks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\acp{CNN} are designed to work with grid-like input structures that exhibit
strong local spatial dependencies. Although most work with \acp{CNN}
involve image-based data, they can be applied to other spatially adjacent data
types such as time-series and text items. \acp{CNN} are defined by the use of a
convolution operation, a mathematical operation that expresses the amount
overlap between functions. Much like traditional neural networks the convolution operation in this context involves multiplying the input by an array of weights, called a filter or a kernel qhich is typically smaller in size than the input. The convolution is applied by shifting the kernel over the input, drawing out spatially important features between the
two. The distance by which the grid is shifted is known as
the stride and increasing it reduces the dimensionality of the
output. The output of the convolutional layer is then passed to an activation function.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GANs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% basic intro to GANs
%
A subset of deep learning that has seen fruitful development in recent years
are \acp{GAN}~\cite{Goodfellow2014}. These unsupervised algorithms learn patterns in a
given training data set using an adversarial process. The generations from
\acp{GAN} are currently state-of-the-art in fields such as high quality image
fidelity~\cite{brock2018large,karras2019analyzing}, text-to-image
translation~\cite{reed2016generative}, and video
prediction~\cite{liang2017dual} as well as time series
generations~\cite{esteban2017realvalued}~\chris{do we still not have any other
GW GAN applications to reference or are we hoping to still be the first?}. \jordan{None that i can find?}

% Basic components of a GAN
%
\acp{GAN} train two competing neural networks, consisting of a discriminator
network that is set up to distinguish between real and fake data and a
generator network that produces fake versions of the real data. The generator model performs a mapping from a fixed length vector $\mathbf{z}$ to its
representation of the data. The input vector is drawn randomly from a Gaussian distribution which is referred to as a latent space comprised of latent variables. The latent space is a compressed representation of a data distribution which the generator applies meaning to during training. Sampling points from this space allows the generator to produce a variety of different generations, with different points corresponding to different features in the generations. The discriminator maps its input $\mathbf{x}$ to a probability that the input came from either the training (real) data or
generator (fake).

\begin{comment}
Chris' definition of latent space if needed: It maps the intrinsic
variation of the training space onto the Gaussian latent space distribution.
\end{comment}
